---
title: "NCT011500045"
author: "Claudia Pascual Tomàs"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tabla de Distribución

<https://gt.rstudio.com/articles/case-study-clinical-tables.html>

```{r}
library(readr)

rx_adsl <- readr::read_csv("NCT01150045_unido.csv")

str(rx_adsl)

colSums(is.na(rx_adsl[, c("age","sex","race","ethnicity", "aspirin", "bmi", "tstage", "nstage_cal", "sidedness", "ps" )]))

```

```{r}
library(rlang)

rx_adsl <- rx_adsl |>
  dplyr::mutate(
    arm = dplyr::case_when(
      grepl("Celecoxib", arm) ~ "Celecoxib",
      grepl("Placebo", arm) ~ "Placebo"
    )
  )
#ordenar manualmente race:
rx_adsl <- rx_adsl |>
  dplyr::mutate(
    race = factor(
      race,
      levels = c(
        "White",
        "Black or African American",
        "Asian",
        "All others or not reported"
      )
    )
  )
custom_summary <- function(df, group_var, sum_var) {
  group_var <- rlang::ensym(group_var)
  sum_var <- rlang::ensym(sum_var)

  is_categorical <-
    is.character(df[[as_string(sum_var)]]) |
    is.factor(df[[as_string(sum_var)]])

  if (is_categorical) {
    category_lbl <-
      as_string(sum_var)
#para quitar los NA de la tabla. modificado con chatgpt
    df_out <-
      df |>
      dplyr::filter(!is.na(!!sum_var)) |> #chatgpt
      dplyr::group_by(!!group_var) |>
      dplyr::mutate(den = dplyr::n()) |> #chatgpt
      dplyr::group_by(!!group_var, !!sum_var) |>
      dplyr::summarize(
        val = dplyr::n(),
        den = unique(den),
        pct = val / den,
        .groups = "drop"
      ) |>
      tidyr::pivot_wider(
        id_cols = !!sum_var, names_from = !!group_var,
        values_from = c(val, den, pct)
      ) |>
      dplyr::rename(label = !!sum_var) |>
      dplyr::mutate(
        across(where(is.numeric), ~ ifelse(is.na(.), 0, .)),
        category = category_lbl
      )
  } else {
    category_lbl <- as_string(sum_var)

    df_out <-
      df |>
      dplyr::group_by(!!group_var) |>
      dplyr::summarize(
        n = sum(!is.na(!!sum_var)),
        mean = mean(!!sum_var, na.rm = TRUE),
        sd = sd(!!sum_var, na.rm = TRUE),
        median = median(!!sum_var, na.rm = TRUE),
        lcl = min(!!sum_var, na.rm = TRUE),
        ucl = max(!!sum_var, na.rm = TRUE),
        .groups = "drop"
      ) |>
      tidyr::pivot_longer(
        cols = c(mean, median), #quito n, no hace falta
        names_to = "label",
        values_to = "val"
      ) |>
      dplyr::mutate(
        sd = ifelse(label == "mean", sd, NA),
        label = dplyr::recode(
          label,
          "mean" = "Mean (SD)",
          "median" = "Median (Range)"
        )
      ) |>
      tidyr::pivot_wider(
        id_cols = label,
        names_from = !!group_var,
        values_from = c(val, sd, lcl, ucl)
      ) |>
      dplyr::mutate(category = category_lbl)
  }

  return(df_out)
}

adsl_summary <-
  dplyr::filter(rx_adsl, primary_analysis == "Yes") |> #cambiamos rnadomized por primany_analyisis
  (\(data) purrr::map_df(
    .x = dplyr::vars(age, sex, race, ethnicity, aspirin, bmi, ps, tstage, nstage_cal, sidedness),
    .f = \(x) custom_summary(df = data, group_var = arm, sum_var = !!x)
  ))()
#esto lo ha hecho chatgpt:
adsl_summary <- adsl_summary |>
  dplyr::mutate(
    category = dplyr::recode(
      category,
      age = "Age, y",
      sex = "Sex",
      race = "Race",
      ethnicity = "Hispanic or Latino",
      bmi = "BMI",
      aspirin = "Aspirin",
      ps = "ECOG performance Status",
      tstage = "Extent of invasion through bowel wall, No./total (%)",
      nstage_cal = "Nodal Stage",
      sidedness = "Tumor location, No./total. (%)"
    )
  )
adsl_summary <- adsl_summary |>
  dplyr::filter(!(category == "Aspirin" & label == "No")) |> #borramos la fila no
  dplyr::mutate(
    label = dplyr::case_when(
      category == "Aspirin" & label == "Yes" ~ "Low dose aspirin usage",
      TRUE ~ label #cambio la etiqueta Yes por Low dose..
    )
  )

adsl_summary <- adsl_summary |>
  dplyr::mutate(
    category = dplyr::case_when(
      category == "Aspirin" ~ "",
      TRUE ~ category
    )
  )

adsl_summary <- adsl_summary |>
  dplyr::mutate(
    label = dplyr::case_when(
      category == "BMI" & label == "Median (Range)" ~ "Median",
      TRUE ~ label,
    )
  ) 

```

```{r}
library(gt)

rx_adsl_tbl <-
  adsl_summary |>
  gt(
    rowname_col = "label",
    groupname_col = "category"
  ) |>
  tab_header(
    title = "Baseline Patient Characteristics in Primary Analysis Population" )

rx_adsl_tbl <-
  rx_adsl_tbl |>
  fmt_integer(
    columns = starts_with(c("val_")),
    rows = !(label %in% c("Mean (SD)", "Median")) #para que sean todos los valores enteros excepto media y mediana
  ) |>
  fmt_percent(
    columns = starts_with("pct_"),
    decimals = 1,
  ) |>
  fmt_number(
    columns = starts_with("val_"),
    rows = label == "Mean (SD)",
    decimals = 1
  ) |>
  fmt_number(
    columns = starts_with("sd_"),
    rows = label == "Mean (SD)",
    decimals = 2
  ) |>
  fmt_number(
    columns = starts_with("lcl_"),
    decimals = 1
  ) |>
  fmt_number(
    columns = starts_with("ucl_"),
    decimals = 1
  )|>
  fmt_number(
  columns = starts_with("val_"),
  rows = label == "Median (Range)",
  decimals = 1
  ) |>
  fmt_number(
  columns = starts_with("val_"),
  rows = label == "Median",
  decimals = 1
  )

rx_adsl_tbl <- rx_adsl_tbl |>
  cols_merge(
    columns = c("val_Placebo", "lcl_Placebo", "ucl_Placebo"),
    rows = label == "Median (Range)" & category == "Age, y",
    pattern = "{1} ({2}–{3})"
  ) |>
  cols_merge(
    columns = c("val_Celecoxib", "lcl_Celecoxib", "ucl_Celecoxib"),
    rows = label == "Median (Range)" & category == "Age, y",
    pattern = "{1} ({2}–{3})"
  ) 
rx_adsl_tbl <-
  rx_adsl_tbl |>
  cols_merge(
    columns = c("val_Placebo", "pct_Placebo", "sd_Placebo"),
    rows = !category %in% c(
      "Extent of invasion through bowel wall, No./total (%)",
      "Tumor location"
    ),
    pattern = "<<{1}>><< ({2})>><< ({3})>>"
  ) |>
  cols_merge(
    columns = c("val_Celecoxib", "pct_Celecoxib", "sd_Celecoxib"),
    rows = !category %in% c(
      "Extent of invasion through bowel wall, No./total (%)",
      "Tumor location"
    ),
    pattern = "<<{1}>><< ({2})>><< ({3})>>"
  )
rx_adsl_tbl <- rx_adsl_tbl |>
  cols_merge(
    columns = c("val_Placebo", "den_Placebo", "pct_Placebo"),
    rows = category %in% c(
      "Extent of invasion through bowel wall, No./total (%)",
      "Tumor location"
    ),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    columns = c("val_Celecoxib", "den_Celecoxib", "pct_Celecoxib"),
    rows = category %in% c(
      "Extent of invasion through bowel wall, No./total (%)",
      "Tumor location"
    ),
    pattern = "{1}/{2} ({3})"
  )

rx_adsl_tbl <-
  rx_adsl_tbl |>
  tab_stub_indent(
    rows = everything(),
    indent = 5
  ) |>
  opt_align_table_header(align = "left")
rx_adsl_tbl <-
  rx_adsl_tbl |>
  cols_width(
    starts_with("val_") ~ px(200),
    1 ~ px(250)
  ) |>
  cols_align(
    align = "center",
    columns = starts_with("val_")
  )

### Count subjects per arm and summarize values in a list
arm_n <-
  rx_adsl |>
  dplyr::filter(primary_analysis == "Yes") |>
  dplyr::group_by(arm) |>
  dplyr::summarize(
    lbl = sprintf("%s N=%i (100%%)", unique(arm), dplyr::n()),
    .groups = "drop"
  ) |>
  dplyr::arrange(arm)

collbl_list <- as.list(arm_n$lbl)
names(collbl_list) <- paste0("val_", arm_n$arm)

rx_adsl_tbl <-
  rx_adsl_tbl |>
  cols_label(.list = collbl_list)

rx_adsl_tbl

```

<https://gt.rstudio.com/reference/gtsave.html>

```{r}
#rx_adsl_tbl |> gtsave("rx_adsl_tbl.png", expand = 10)
#rx_adsl_tbl |> gtsave(filename = "rx_adsl_tbl.html", inline_css = TRUE)
```

# Survival Analysis in R

https://rpubs.com/mbounthavong/survival_analysis_in_r 

## Kaplan Meier

```{r}
names(rx_adsl)
```


```{r}
library(survival)

table(rx_adsl$os_stat)

table(rx_adsl$dfs_stat)

rx_adsl <- rx_adsl |> #337-347 chatgpt
  dplyr::mutate(
    os_event = dplyr::case_when(
      os_stat == "1=Event" ~ 1,
      os_stat == "0=Censor" ~ 0
    ),
    dfs_event = dplyr::case_when(
      os_stat == "1=Event" ~ 1,
      os_stat == "0=Censor" ~ 0
  )
)

kmcurve_os <- survfit(Surv(os_time, os_event) ~ arm, data = rx_adsl)
plot(kmcurve_os)
kmcurve_dfs <- survfit(Surv(dfs_time, dfs_event) ~ arm, data = rx_adsl)
plot(kmcurve_dfs)
```
```{r}
library(survminer)
ggsurvplot(kmcurve_os, 
           conf.int = TRUE,
           pval = TRUE,
           risk.table = TRUE,
           legend.labs = c("Celecoxib", "Placebo"),
           title = "OS Kaplan-Meier curve")
ggsurvplot(kmcurve_dfs, 
           conf.int = TRUE,
           pval = TRUE,
           risk.table = TRUE,
           legend.labs = c("Celecoxib", "Placebo"),
           title = "DFS Kaplan-Meier curve")
```
```{r}
survdiff(Surv(os_time, os_event) ~ arm, data = rx_adsl)

survdiff(Surv(dfs_time, dfs_event) ~ arm, data = rx_adsl)

```
Both the ggsurvplot and survdiff yield a p-value for the log rank test > 0,05. Therefore, we fail to reject the null that there is no statistically significant difference in survival between Celecoxib and Placebo.


## Cox proportional hazard model 

###Overall survival:

```{r}

coxmodel_os <- coxph(Surv(os_time, os_event) ~ arm, data = rx_adsl)

exp(coef(coxmodel_os))

exp(confint(coxmodel_os))
```
```{r}
library(gtsummary)
model1 <- tbl_regression(coxmodel_os, intercept = TRUE)
as_gt(model1) %>% 
              gt::tab_header("Cox proportional hazards regression model") %>% 
              gt::tab_options(table.align='left')
```
The hazard ratio was 1.13 with a 95% confidence interval of 0.95, 1.37. Since the 95% confidence interval crosses where the HR=1, we fail to reject the null hypothesis that there is no significant differences in hazards between subjects in Treatment with Celecoxib and Placebo.

###Disease free survival:

```{r}
coxmodel_dfs <- coxph(Surv(dfs_time, dfs_event) ~ arm, data = rx_adsl)

exp(coef(coxmodel_dfs))

exp(confint(coxmodel_dfs))
```
```{r}
library(gtsummary)
modeldfs <- tbl_regression(coxmodel_dfs, intercept = TRUE)
as_gt(modeldfs) %>% 
              gt::tab_header("Cox proportional hazards regression model") %>% 
              gt::tab_options(table.align='left')
```
The hazard ratio was 1.16 with a 95% confidence interval of 0.97, 1.40. Since the 95% confidence interval crosses where the HR=1, we fail to reject the null hypothesis that there is no significant differences in hazards between subjects in Treatment with Celecoxib and Placebo.
